---
title: "Examples for ddst package"
author: "PB"
date: "July 26, 2019"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown

This document presents showcases for the test avaliable in the ddst package.

```{r, warning=FALSE, message=FALSE}
library("ddst")
```

# II. Data driven smooth tests for selected goodness-of-fit problems

```{r}
```

## 1. Data Driven Smooth Test for Uniformity

```{r}
z <- c(13.41, 6.04, 1.26, 3.67, -4.54, 2.92, 0.44, 12.93, 6.77, 10.09,
       4.10, 4.04, -1.97, 2.17, -5.38, -7.30, 4.75, 5.63, 8.84)
ddst.uniform.test(z, compute.p = TRUE)

# when H0 is true
z <- runif(80)
ddst.uniform.test(z, compute.p = TRUE)

# for known fixed alternative, here N(10,16)
z <- pnorm(z, 10, 16)
ddst.uniform.test(z, compute.p = TRUE)

# when H0 is false
z <- rbeta(80,4,2)
ddst.uniform.test(z, compute.p = TRUE)

# just p-value
ddst.uniform.test(z, compute.p = TRUE)$p.value
```
## 2. Data Driven Smooth Test for Exponentiality

```{r}
# for given vector of 19 numbers
z <- c(13.41, 6.04, 1.26, 3.67, -4.54, 2.92, 0.44, 12.93, 6.77, 10.09,
       4.10, 4.04, -1.97, 2.17, -5.38, -7.30, 4.75, 5.63, 8.84)
ddst.exp.test(z, compute.p = TRUE)

# when H0 is true
z <- rexp(80, 4)
ddst.exp.test(z, compute.p = TRUE)

# when H0 is false
z <- rchisq(80, 4)
ddst.exp.test(z, compute.p = TRUE)

```

## 3. Data Driven Smooth Tests for Normality


### 3.1. The case of bounded basis functions

```{r}
# for given vector of 19 numbers
z <- c(13.41, 6.04, 1.26, 3.67, -4.54, 2.92, 0.44, 12.93, 6.77, 10.09,
      4.10, 4.04, -1.97, 2.17, -5.38, -7.30, 4.75, 5.63, 8.84)
ddst.norm.test(z, compute.p = TRUE)

# when H0 is true
z <- rnorm(80)
ddst.norm.test(z, compute.p = TRUE)

# when H0 is false
z = rexp(80, 4)
ddst.norm.test(z, B = 5000, compute.p = TRUE)

```

### 3.2. The case of unbounded basis functions



## 4. Data Driven Smooth Test for Extreme Value Distribution


```{r}
# for given vector of 19 numbers
z <- c(13.41, 6.04, 1.26, 3.67, -4.54, 2.92, 0.44, 12.93, 6.77, 10.09,
      4.10, 4.04, -1.97, 2.17, -5.38, -7.30, 4.75, 5.63, 8.84)
ddst.extr.test(z, compute.p = TRUE)

library(evd)
z <- -qgumbel(runif(100), -1, 1)
ddst.extr.test(z, compute.p = TRUE)

# when H0 is false
z <- rexp(80, 4)
ddst.extr.test(z, compute.p = TRUE)

```

# III. Nonparametric Data Driven Smooth Tests for Comparing Distributions


## 5. Data Driven Smooth Test for k-Sample Problem


## 6. Data Driven Smooth Test for Stochastic Dominance in Two Samples


## 7. Two-Sample Test Against Stochastic Dominance

```{r}
# H0 is true
x = runif(80)
y = runif(80)
ddst.twosample.test(x, y, compute.p=TRUE)

# known fixed alternative
x = runif(80)
y = rbeta(80,4,2)
ddst.twosample.test(x, y, compute.p=TRUE)
```

## 8. Data Driven Smooth Test for Upward Trend Alternative in k samples


## 9. Data Driven Smooth Test for Umbrella Alternatives in k samples


## 10. Data Driven Smooth Test for Change-Point Problem


## Two-Sample Test Against One-Sided Alternatives

